---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv", header = TRUE)
```

## What is mean total number of steps taken per day?
<!-- 
# Using all values for steps per day including NAs
stepsperday <- sapply(split(activity$steps, activity$date), sum)
# Histogram will show NAs as zero which is not necessarily what we want
hist(stepsperday)
# mean() and median() uses unmodified length but ignores na's
mean(stepsperday, na.rm = TRUE)
median(stepsperday, na.rm = TRUE)
-->

<!--
# Removing all lines where values are missing
# Note that this removes the NAs completely, reducing length() used in mean() and median()
complete_activity <- activity[!is.na(activity$steps),]
# Creating a labeled integer vector of total steps per day 
stepsperday <- sapply(split(complete_activity$steps, complete_activity$date), sum)
# Histogram completely ignores NA's as they are not present in this data set
hist(stepsperday)
# mean() and median() uses the length of modified data set without NAs, which might not be want we wanted
mean(stepsperday)
median(stepsperday)
-->

```{r}
# Creating a named integer vector of total steps per day 
stepsperday <- sapply(split(activity$steps, activity$date), sum)
```
Our named integer vector will contain NA values just like the original data set. The histogram will ignore these NA values.
```{r}
hist(stepsperday)
```
The mean and median function can not handle na-values so these are removed. Removing them using the na.rm flag will read then as 0 but use the same length as the original list. Creating a subset of complete values would give different mean and median values. 
```{r}
mean(stepsperday, na.rm = TRUE)
median(stepsperday, na.rm = TRUE)
```

## What is the average daily activity pattern?

```{r}
# Average steps per interval
meanstepsperinterval <- sapply(split(activity$steps, activity$interval), mean, na.rm = TRUE)
# Plot average for each day
plot(meanstepsperinterval, type = "l")
```
Since the integer ector is named with interval we can et the interval name with names().
```{r}
# Interval with highest average number of steps
names(meanstepsperinterval)[meanstepsperinterval == max(meanstepsperinterval)]
```

## Inputing missing values

```{r}
# Total number of missing values in activity data set
sum(is.na(activity))
# Replacing NA with mean for that interval
filledactivity <- activity
for(i in 1:nrow(filledactivity)) {
	if(is.na(filledactivity$steps[i])) {
		interval <- as.character(filledactivity$interval[i])
		meansteps <- meanstepsperinterval[interval]
		filledactivity$steps[i] <- meansteps
	}
}
```

```{r}
filledstepsperday <- sapply(split(filledactivity$steps, filledactivity$date), sum)
hist(filledstepsperday)
mean(filledstepsperday, na.rm = TRUE)
median(filledstepsperday, na.rm = TRUE)
```
Since I filled NA values with the interval average we clearly see that the center of our histogram is stretched taller with a higher frequence of average values. Our mean and median converge to the same value.

## Are there differences in activity patterns between weekdays and weekends?
We can replace our old date with a Date object.

```{r}
filledactivity$date <- as.POSIXlt(filledactivity$date, format = "%Y-%m-%d")
```

```{r results="hide"}
# As I am using a Swedish version of OSX I need to set my locale to english in order to use 'Saturday' and 'Sunday' as my weekday-tests.
Sys.setlocale(locale = "en_US.UTF-8")
```
```{r}
filledactivity$weekday <- factor(ifelse(weekdays(filledactivity$date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
weekdayactivity <- filledactivity[filledactivity$weekday == "weekday",]
meanweekday <- sapply(split(weekdayactivity$steps, weekdayactivity$interval), mean)
weekendactivity <- filledactivity[filledactivity$weekday == "weekend",]
meanweekend <- sapply(split(weekendactivity$steps, weekendactivity$interval), mean)
par(mfrow = c(2,1))
plot(meanweekday, type = "l", ylim = c(0, 250), xlim = c(0, 300), ylab = "", xlab = "Interval", main = "Weekdays")
plot(meanweekend, type = "l", ylim = c(0, 250), xlim = c(0, 300), ylab = "", xlab = "Interval", main = "Weekends")
par(mfrow = c(1,1), mar = c(2,2,2,2))
mtext(text = "Number of steps", side = 2)
```
```{r}
```
```{r}
```
